### 2010-09-15 notes/proposal.tex
Work on proposal

\documentclass[11pt]{article}

\input{format.tex}

\title{M.Sc Proposal}
\author{Roman Zeyde}
\begin{document}
\maketitle
\section{Introduction}
Electrophoresis is the motion of dispersed particle relative to a ion-containing fluid,
due to the influence of an electric field. This phenomenon combines electrostatics
coupled with fluid dynamics, as well as diffusion/advection of ions in the fluid.

The particle (surrounded by fluid) acquire surface charge which is screened by
a thin layer of ions in the fluid, creating a double-layer structure near the
particle's surface, called ``Debye layer''.

If the layer width is much smaller than the particle's size
an analytical solution in the Debye-layer can be found \cite{Yariv10},
and be used as the appropriate boundary condition for the
macroscale problem on the particle's surface.

Electrophoresis is described by the following physical phenomena,
using the macroscale equations:
\begin{enumerate}
\item Electrostatic potential - the fluid contains no free charge:
\begin{equation}
    \bnabla \cdot (C \bnabla \varPhi) = 0.
\end{equation}
\item Fluid dynamics - Stokes incompressible flow with electrostatic force component:
\begin{equation}
\left\{ \begin{array}{l}
\bLaplacian \bV - \bnabla P + \Laplacian \varPhi \bnabla \varPhi = 0 \\
\bnabla \cdot \bV = 0 \end{array} \right.
\end{equation}
\item Ions' diffusion and advection - through the fluid's velocity field:
\begin{equation}
\Laplacian C - \alpha \bV \cdot \bnabla C = 0
\end{equation}
\end{enumerate}

The boundary conditions correspond to the specific problem at hand, and
are coupled as well as the partial differential equations themselves.

The closed form approximate solution was developed for spherical particle
and small electric field \cite{Yariv10} but since the equations are coupled
and non-linear, it is hard to extend the analytic solution to more general
settings.

\section{Research goals}
Our goal is to implement an accurate and fast iterative numerical
solver for the problem described above,
in order to enable the research of cases which has no closed form solution
(e.g. non-spherical/asymmetric particles, large electric fields).

The solver will be verified against known closed-form solutions
for each of the sub-problems, as well as the full coupled problem,
analyzing the discrepancies between the theoretical and numerical results.

Moreover, the convergence rate of the solver will be optimized using
numerical acceleration methods (e.g. MPE, RRE and Multigrid) to achieve
high accuracy in reasonable solver's execution time.

Achieved results will provide a numeric model for interesting small-scale physical phenomena,
such as an ``self-electrophoretic swimmer'' by numerically modeling an asymmetric particle's configuration.

\section{Achieved results}
The numerical solver is being implemented as a MATLAB package that
provides the needed tools for numerical solution of an electrophoresis problem.

The solver applies iterative relaxation to solve the system, by
iterating the equations and solving repeatedly each one of them
for the corresponding variable
(Laplace equation for $\varPhi$, Stokes equation for $(\bV, P)$
and diffusion-advection equation for $C$), where
all other variables keep their values from previous iteration
(a ``Gauss-Seidel'' method for non-linear systems).

Problem's configuration is taken from \cite{Yariv10} and
consists of a spherical particle, surrounded by infinite Stokes fluid.
Suppose that the particle is stationary and fluid flows around it, slipping across
the boundary layer (due to ``Debye layer'' effects) with a velocity field denoted
by $\bV$.

In this setting, together with the usage of spherical coordinates $(R,\theta,\phi)$,
makes the problem axisymmetric around the velocity vector of the fluid (independent
of $\phi$). Thus the problem may be written as two-dimensional partial
differential equation system in $R$ and $\theta$ variables,
such that $R \in [1,\infty)$ and $\theta \in [0, \pi]$.
A natural choice of a grid is logarithmic grid for $R$ and uniform grid for $\theta$.

The following linear operators are implemented in spherical coordinates using regular grid:
\begin{itemize}
  \item Scalar Laplacian (for $\varPhi$ and $C$)
  \item Vector Laplacian (for $\bV$)
  \item Gradient (for $\Phi, C$ and $P$)
  \item Divergence (for $\bV$)
\end{itemize}

Thus, each differential operator is written in spherical coordinates on the regular
$(r, \theta)$ grid and being linear -- is represented as a sparse matrix,
making the algebraic computations on the equations much more straightforward.

The boundary conditions are computed for $\theta = 0$ and $\theta = \pi$ using symmetry
considerations. At $R = \infty$, we assume constant fluid velocity
and $\bV = \cal{U} \bzhat$, constant electric field $\bnabla \varPhi = \beta \bzhat$
and constant ionic concentration $C = 1$.
The boundary conditions on $R = 1$ are defined by the boundary layer behavior as described
in \cite{Yariv10}.

After substitution of boundary conditions, each differential equation is transformed to a
sparse linear system of the form $\matr{A}\bx = \vect{b}$.
Due to coupling, it is not necessary to find the exact solution -- so we apply
few relaxations on each linear system, using an appropriate preconditioner $\matr{M}$
(which depends on $\matr{A}$):
\begin{eqnarray}
  \bx_{n+1} &=& \bx_{n} + \matr{M}_\matr{A}^{-1} \left(\vect{b} - \matr{A}\bx_{n}\right)
\end{eqnarray}

After the solver has converged to a solution, the stress tensor $\mathbb{T}$ on the particle
$\cal{S} = \left\{\bx : \|\bx\| = 1\right\}$ is computed:
\begin{eqnarray}
  \vect{F} = \int_{\cal{S}} \left(\mathbb{T} \cdot \bnhat \right) ds = \vec{0}.
\end{eqnarray}
Since the problem is axisymmetric, the force should be aligned with the symmetry axis:
$\vect{F} = F \bzhat$. Moreover, since the particle is stress-free in steady flow, we
have $F = 0$ for steady state velocity $\cal{U}$.

This observation enables us to find the correct $\cal{U}$ for given $\beta$, using one-dimensional root-finding algorithm on $F_\beta(\cal{U})$ function.

\section{Algorithms and techniques}

\subsection{Grid choice}
$C$, $\varPhi$ and $P$ are using central grid whereas staggered grid is used for $\bV$.

\subsection{Differential Equations}
Finite volume method using finite differences.
Upwind for advection.

\subsection{Preconditioners}
There are efficient sparse implementations of 
Jacobi, Red-Black Gauss-Seidel and Vanka-type.

\begin{thebibliography}{}

\bibitem{Yariv10} E. Yariv,
``Migration of ion-exchange particles driven by a uniform electric field'',
{\em Journal of Fluid Mechanics, 655 105-121}, 2010.

\end{thebibliography}

\end{document}