### 2011-11-27 notes/summary/ch02.tex

\section{Numerics}
\subsection{Discretization}
\subsubsection{Spherical Coordinates}
Since the system has axial symmetry, the operators are written in 
spherical coordinates $(r,\theta,\phi)$:
\begin{eqnarray}
x &=& r \sin \theta \cos \phi \\
y &=& r \sin \theta \sin \phi \\
z &=& r \cos \theta
\end{eqnarray}
Due to axial symmetry, $\deriv{a}{\phi} = 0$ and $\bphihat \cdot \bA = 0$, 
so the differential operators have the following form:
\subsubsection{Grids choice}
\subsubsection{Operator Representation}
\subsubsection{Boundary Condition}
\subsection{Solver}
\subsubsection{Newton's Method}
\subsubsection{Steady-state}
\subsubsection{Continuation}