### 2011-12-02 notes/summary/ch01.tex
Solver docs.

\section{Physics}
Electrokinetic phenomena is described by the equations described below.
An ion-exchanger is described by the chemical kinetics at its boundary.
The experiment consists of applying an electic field and measuring the drift velocity of the particles.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Micro-scale phenomena}
Denote positive and negative ionic concentration by $c_+$ and $c_-$ respectively.
The respective fluxes are denoted by $\bj_+$ and $\bj_-$.

The electrostatic potential is denoted by $\varphi$.

The velocity and the pressure field are denoted by $\bv$ and $p$ respectively.

\subsubsection{Nernst-Planck Equations}
The ionic fluxes are given by ions' diffusion, electrostatic forces and ions' advection by the fluid:
\begin{eqnarray}
  \bj_\pm &=& -\bnabla c_+ \mp c_+ \bnabla \varphi + \alpha \bv c_\pm
\end{eqnarray}
Due to conservation of ions, the fluxes are divergence-free:
\begin{eqnarray}
\bnabla \cdot \bj_\pm &=& 0.
\end{eqnarray}
These equation may be re-written by using the following transform:
\begin{eqnarray}
  c &=& \frac{c_+ + c_-}{2}\\
  q &=& \frac{c_+ - c_-}{2}
\end{eqnarray}
Thus, salt and charge fluxes are defined by:
\begin{eqnarray}
  \bj &=& \frac{\bj_+ + \bj_-}{2} = -\bnabla c - q \bnabla \varphi + \alpha \bv c \\
  \bi &=& \frac{\bj_+ - \bj_-}{2} = -\bnabla q - c \bnabla \varphi + \alpha \bv q
\end{eqnarray}
And are divergence-free as well:
\begin{eqnarray}
\bnabla \cdot \bj &=& 0 \\
\bnabla \cdot \bi &=& 0 
\end{eqnarray}

\subsubsection{Incompressible Stokes Flow with Electrostatic Force}
Mass conservation (due to constant fluid density):
\begin{eqnarray}
\bnabla \cdot \bv &=& 0
\end{eqnarray}
Momentum conservation (force balance):
\begin{eqnarray}
\Laplacian \bv - \bnabla p + \bnabla \varphi \Laplacian \varphi &=& 0
\end{eqnarray}
Free charge density is given by Poisson equation:
\begin{eqnarray}
\delta^2 \Laplacian \varphi &=& -q
\end{eqnarray}

\subsubsection{Boundary conditions}
Away of the particle, we have uniform flow, uniform electrostatic field
and uniform ionic concentrations:
\begin{eqnarray}
\bv &\rightarrow& -\cU \ui \\
\bE = -\bnabla \varphi &\rightarrow& \beta\ui \\
c_\pm &\rightarrow& 1
\end{eqnarray}

We consider a spherical particle. At its surface ($r=1$ with normal $\bn = \brhat$), 
we have zero slip, anion impermeability, cation selectivity and particle's conductance:
\begin{eqnarray}
\bv & = & \bzero \\
\bn \cdot \bj_- & = & 0 \\
c_+ & = & \gamma \\
\varphi & = & \cV
\end{eqnarray}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Separation of scales}
Due to strong electrostatic forces, 
most of the fluid remains electroneutral ($c_+ \approx c_-$), 
except of a thin boundary layer (``Debye layer'') around the ion exchanger.

\subsubsection  {Bulk-scale equations}
Nernst-Planck equations above for the fluid bulk (outside
the boundary layer) using bulk variables:
\begin{eqnarray}
  Q & = & 0 \\
  C & = & C_+ = C_- \\
\bJ &=& -\bnabla C + \alpha \bV C \\
\bI &=& -C \bnabla \varPhi
\end{eqnarray}
Since the flow is incompressible ($\bnabla \cdot V = 0$), 
ion conservation results in the following equations:
\begin{eqnarray}
0 = \bnabla \cdot \bJ &\Rightarrow& \Laplacian C - \alpha \bV \cdot \bnabla C = 0 \\
0 = \bnabla \cdot \bI &\Rightarrow& \bnabla \cdot \pars{ C \bnabla \varPhi } = 0
\end{eqnarray}
Stokes flow equations stay the same:
\begin{eqnarray}
\bnabla \cdot \bV &=& 0 \\
\Laplacian \bV - \bnabla P + \bnabla \varPhi \Laplacian \varPhi &=& 0
\end{eqnarray}

\subsubsection{Debye-scale effective boundary conditions}
We define the rescaled coordinate $\rho$ (normal to ion-exchanger surface), where $\delta \ll 1$:
\begin{eqnarray}
  \rho &=& \frac{r-1}{\delta} 
\end{eqnarray}
Thus, the electric potential and the ionic concentrations are $O(1)$ at Debye layer ($r \approx 1$):
\begin{eqnarray}
  \varphi(\rho,\theta) &\sim& \varphi \\
  c_\pm(\rho,\theta) &\sim& c_\pm \\
  c(\rho,\theta) &\sim& c \\
  q(\rho,\theta) &\sim& q
\end{eqnarray}

The radial fluxes at the Debye layer are $O(\delta^{-1})$, due to coordinate rescaling:
\begin{eqnarray}
  j_\pm(\rho, \theta) &\sim& \delta^{-1} j_\pm
\end{eqnarray}

The pressure is $O(\delta^{-2})$, due to momentum balance equations:
\begin{eqnarray}
  p(\rho, \theta) &\sim& \delta^{-2} p
\end{eqnarray}

The tangential velocity component is $O(1)$ and the radial one is $O(\delta)$.

In order to match bulk $O(1)$ outer radial ionic flux, 
the $O(\delta^{-1})$ inner ionic radial flux $j_\pm(\rho, \theta)$ must vanish:
\begin{eqnarray}
  -\deriv{c_\pm}{\rho} \mp c_\pm \deriv{\varphi}{\rho} &=& 0 \\
  \deriv{\varphi}{\rho} \pm\frac{1}{c_\pm}\deriv{c_\pm}{\rho} &=& 0 \\
  \pars{\varPhi - \varphi} \pm\pars{\log C_\pm - \log c_\pm} &=& 0 \\
  \varPhi - \varphi \pm \log \frac{C_\pm}{c_\pm} &=& 0
\end{eqnarray}
This relation can be rewritten as Boltzmann distribution of ionic concentration,
since $C_\pm|_{\rho\rightarrow\infty} = C$:
\begin{eqnarray}
c_\pm = C \exp\left[\pm(\varPhi - \varphi)\right]
\end{eqnarray}

Due to cation selectivity, $c_+|_{\rho=0} = \gamma$.
\begin{eqnarray}
  c_+ = \gamma &=& C \exp(\varPhi - \cV) \\
  \cV + \log \gamma &=& \log C + \varPhi
\end{eqnarray}

Denote zeta potential as the voltage drop between the particle surface and the end of
the boundary layer, $\zeta = \cV - \varPhi$. 
Without loss of generality, 
we assume that $\cV = -\log\gamma$, thus $\zeta = -\log\gamma-\varPhi$:
\begin{eqnarray}
\varPhi &=& -\log C
\end{eqnarray}

Due to anion impermeability and radial flux continuity:
\begin{eqnarray}
  \br \cdot \bj_- = -\deriv{C_-}{r} + C_- \deriv{\varPhi}{r} &=& 0 \\
-\deriv{C}{r} + C \deriv{\varPhi}{r} &=& 0
\end{eqnarray}

The leading-order electric potential in the Debye layer:
\begin{eqnarray}  
\deriv{^2 \varphi}{\rho^2} = -q = -\frac{c_+ - c_-}{2} = 
-C\frac{\exp\left[+(\varPhi - \varphi)\right] - \exp\left[-(\varPhi - \varphi)\right]}{2} = 
C \sinh(\varphi - \varPhi)
\end{eqnarray}
Denote $\psi = \varphi - \varPhi$, such that $\psi(0) = \zeta$ and 
$\psi(\rho\rightarrow\infty) = 0$.
\begin{eqnarray}  
\deriv{^2 \psi}{\rho^2} &=& C \sinh(\psi) = 2 C \sinh{\frac{\psi}{2}}\cosh{\frac{\psi}{2}} 
\\
2\deriv{\psi}{\rho} \deriv{^2 \psi}{\rho^2} &=& 
4 C \sinh{\frac{\psi}{2}}\cosh{\frac{\psi}{2}} \deriv{\psi}{\rho} 
\\
\deriv{}{\rho} \pars{\deriv{\psi}{\rho}}^2 &=& 
\deriv{}{\rho} \pars{2 \sqrt{C} \sinh{\frac{\psi}{2}}}^2 
\end{eqnarray}

Integrating with respect to boundary condtions yields:
\begin{eqnarray}  
\deriv{\psi}{\rho} &=& -2\sqrt{C}\sinh\frac{\psi}{2} \\  
\pars{4\sinh\frac{\psi}{4}\cosh\frac{\psi}{4}}^{-1}\deriv{\psi}{\rho} &=& -\sqrt{C} \\  
\frac{\pars{\tanh\frac{\psi}{4}}^{-1}}{4\cosh^2\frac{\psi}{4}}
\deriv{\psi}{\rho} &=& -\sqrt{C} \\  
\deriv{}{\rho}\pars{\log\tanh\frac{\psi}{4}} &=& \deriv{}{\rho}\pars{-\rho \sqrt{C}} \\
\log\tanh\frac{\psi}{4} - \log\tanh\frac{\zeta}{4} &=& -\rho \sqrt{C} \\
\psi &=& 4 \tanh^{-1}\pars{\tanh{\frac{\zeta}{4}} \cdot \exp\pars{-\rho \sqrt{C}}}
\end{eqnarray}

Leading-order radial momentum balance in the Debye layer yields:
\begin{eqnarray} 
- \deriv{p}{\rho} + \deriv{\varphi}{\rho} \deriv{^2\varphi}{\rho^2} &=& 0
\end{eqnarray}
One integration yields:
\begin{eqnarray} 
p &=& \frac{1}{2}\pars{\deriv{\varphi}{\rho}}^2
\end{eqnarray}

Leading-order tangential momentum balance in the Debye layer yields:
\begin{eqnarray} 
0 &=& \deriv{^2 v_\theta}{\rho^2} - \deriv{p}{\theta} 
 + \deriv{\varphi}{\theta} \deriv{^2\varphi}{\rho^2} 
\\
\deriv{^2 v_\theta}{\rho^2} &=& \frac{1}{2}\deriv{}{\theta}\pars{\deriv{\psi}{\rho}}^2
- \deriv{(\varPhi + \psi)}{\theta} \deriv{^2\psi}{\rho^2}  
\\
 &=& \deriv{}{\theta}\pars{2C \sinh^2\frac{\psi}{2}}
- \deriv{\psi}{\theta} C \sinh\psi
- \deriv{\varPhi}{\theta} \deriv{^2\psi}{\rho^2}  
\\
 &=& \deriv{}{\theta}\pars{C \pars{\cosh\psi - 1}}
- C \sinh\psi \deriv{\psi}{\theta}
- \deriv{\varPhi}{\theta} \deriv{^2\psi}{\rho^2}  
\\
 &=& \deriv{C}{\theta} \pars{\cosh\psi - 1}
- \deriv{\varPhi}{\theta} \deriv{^2\psi}{\rho^2}  
 = \deriv{C}{\theta} \pars{2 \sinh^2 \frac{\psi}{2}}
- \deriv{\varPhi}{\theta} \deriv{^2\psi}{\rho^2}  
\\
 &=& -\deriv{C}{\theta} \frac{2}{\sqrt{C}} \pars{\half \sinh \frac{\psi}{2} \deriv{\psi}{\rho}}
- \deriv{\varPhi}{\theta} \deriv{^2\psi}{\rho^2}  
 = -\deriv{C}{\theta} \frac{2}{\sqrt{C}} \deriv{}{\rho}\pars{\cosh \frac{\psi}{2}}
- \deriv{\varPhi}{\theta} \deriv{^2\psi}{\rho^2}  
\end{eqnarray}
By integrating from specific $\rho$ to $\rho \rightarrow \infty$:
\begin{eqnarray}
\deriv{v_\theta}{\rho} &=& -\deriv{C}{\theta} \frac{2}{\sqrt{C}} 
\pars{\cosh \frac{\psi}{2} - 1} - \deriv{\varPhi}{\theta} \deriv{\psi}{\rho}   
= -\deriv{C}{\theta} \frac{4}{\sqrt{C}} \sinh^2 \frac{\psi}{4} 
  - \deriv{\varPhi}{\theta} \deriv{\psi}{\rho}   
\end{eqnarray}
Note that:
\begin{eqnarray}
\deriv{\psi}{\rho} &=& -2\sqrt{C}\sinh\frac{\psi}{2} = 
                       -4\sqrt{C}\sinh\frac{\psi}{4}\cosh\frac{\psi}{4}
\\
-4\sinh\frac{\psi}{4} &=& \frac{1}{\sqrt{C} \cosh\frac{\psi}{4}} \deriv{\psi}{\rho}
\end{eqnarray}
Thus:
\begin{eqnarray}
\deriv{v_\theta}{\rho} &=& 
  \pars{\deriv{C}{\theta} \frac{1}{C}} \cdot
  \pars{\frac{\sinh\frac{\psi}{4}}{\cosh\frac{\psi}{4}} \deriv{\psi}{\rho}}
  - \deriv{\varPhi}{\theta} \deriv{\psi}{\rho}
=  4\deriv{}{\rho} \pars{\log\cosh\frac{\psi}{4}} \deriv{}{\theta} \log C
  - \deriv{\psi}{\rho} \deriv{\varPhi}{\theta}
\end{eqnarray}

By integrating from $\rho = 0$ to $\rho \rightarrow \infty$, 
we have Dukhin-Derjaguin slip formula for tangential velocity component $V_\theta$ 
(where the radial component is $V_r = 0$):
\begin{eqnarray} 
V_\theta &=& \zeta \cdot \deriv{\varPhi}{\theta} -
      4\log\cosh \frac{\zeta}{4} \cdot \deriv{}{\theta} \log C 
\\
&=& \zeta \cdot \deriv{\varPhi}{\theta} + 
      2\log\pars{1 - \tanh^2 \frac{\zeta}{4}} \cdot \deriv{}{\theta} \log C 
\end{eqnarray}
where $\zeta = \cV - \varPhi$.

Since $V_r = 0$ on the boundary, the slip condition can be written using surface
gradient operator:
\begin{eqnarray}
\bV &=& 
\zeta \cdot \bnabla_\mathcal{S} \varPhi 
+ 2\log\pars{1-\tanh^2\frac{\zeta}{4}} \cdot \bnabla_\mathcal{S} \log C \\
\bnabla_\mathcal{S} &=& \pars{\tI - \bn\bn} \bnabla
\end{eqnarray}

\subsection{Steady state}
The stress tensor is composed of Newtonian and Maxwel stresses:
\begin{eqnarray}
\tT &=& \bnabla \bV + (\bnabla \bV)^\dagger - P \tI
+ \bnabla \varPhi \bnabla \varPhi - \frac{1}{2} (\bnabla \varPhi \cdot \bnabla \varPhi) \tI \\
\bnabla \cdot \tT &=& \bzero
\end{eqnarray}
The total force acting on the particle is computed by the following surface integral:
\begin{eqnarray}
\bF &=& \oint_\mathcal{S} \tT \cdot \bnhat  dA 
\end{eqnarray}
In steady state (constant particle drift velocity $\cU$), it should be zero: $\bF = \bzero$.