### 2012-02-21 notes/work/ch04.tex
Review.

\section{Numerical Results}
The results for steady-state velocity $\cU$ as a function of the electric field magnitude $\beta$
are shown at Figures \ref{fig:NumResA} and \ref{fig:NumResB}. 
Since log-log scale is used, blue and red colors are used 
for positive and negative velocities, respectively.

For small $\beta$, the linear regime dominates, $\cU \propto \beta$. 
However, from a specific $\beta_c$, the nonlinear regime dominates, such that
$\cU \propto \beta^3$, until $\beta \sim O(1)$.

The streamlines for a specific $\gamma$ are shown at Figure \ref{fig:NumResC} for
various electric field magnitudes $\beta$.

The theoretical results for the nonlinear regime are computed by asymptotic expansion of the
nonlinear system, as described in the previous section.
\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figs/A1.eps}
        \includegraphics[width=0.45\textwidth]{figs/A2.eps}
        \includegraphics[width=0.45\textwidth]{figs/A3.eps}
        \includegraphics[width=0.45\textwidth]{figs/A4.eps}
        \caption{Steady-state velocity $\cU$ as a function of the electric field magnitude $\beta$, 
        for $\gamma > 1$.}
        \label{fig:NumResA}
    \end{center}
\end{framed}
\end{figure}
\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figs/B1.eps}
        \includegraphics[width=0.45\textwidth]{figs/B2.eps}
        \includegraphics[width=0.45\textwidth]{figs/B3.eps}
        \includegraphics[width=0.45\textwidth]{figs/B4.eps}
        \caption{Steady-state velocity $\cU$ as a function of the electric field magnitude $\beta$,
        for $\gamma < 1$.}
        \label{fig:NumResB}
    \end{center}
\end{framed}
\end{figure}
\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.3\textwidth]{figs/D1n.eps}
        \includegraphics[width=0.3\textwidth]{figs/D1a.eps}

        \includegraphics[width=0.3\textwidth]{figs/D2n.eps}
        \includegraphics[width=0.3\textwidth]{figs/D2a.eps}

        \includegraphics[width=0.3\textwidth]{figs/D3n.eps}
        \includegraphics[width=0.3\textwidth]{figs/D3a.eps}

        \includegraphics[width=0.3\textwidth]{figs/D4n.eps}
        \includegraphics[width=0.3\textwidth]{figs/D4a.eps}
        \caption{Streamlines of the flow for various electric field values: numerical results (left)
        and analytical approximation (right).}
        \label{fig:NumResC}
    \end{center}
\end{framed}
\end{figure}