### 2012-02-21 notes/article/main.tex
Add content to elsarticle.

\documentclass[final]{elsarticle}

\usepackage{graphicx}
\usepackage{amsmath} % Required for \eqref
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{units}   % Required for \nicefrac
\usepackage{float}   % Required for algorithm and floating environments
\usepackage{graphicx}% Required for figures and imagesf
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}
\usepackage{framed}

\floatstyle{ruled}
\newfloat{program}{thp}{lop}
%-------------------------------------------
\newcommand{\sign}{\ensuremath{\mathrm{sign}}}
\newcommand{\diag}{\ensuremath{\mathrm{diag}}}
\newcommand{\trace}[1]{\ensuremath{\mathrm{trace}\left( #1 \right)}}
\newcommand{\norm}[1]{\ensuremath{\left\|#1\right\|_2^2}}
\newcommand{\func}[2]{\ensuremath{\mathrm{#1}\left( #2 \right)}}
\newcommand\eps \epsilon
\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\real}{\ensuremath{\mathbb{R}}}
\newcommand{\cl}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\suppsize}[1]{\ensuremath{|\mathcal{#1}|}}
\newcommand{\vect}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\matr}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\deriv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\arr}[2]{\begin{array}{#1}#2\end{array}}
\newcommand{\mat}[2]{\left(\begin{array}{#1}#2\end{array}\right)}
\newcommand{\brc}[2]{\left\{\begin{array}{#1}#2\end{array}\right.}
\newcommand{\pars}[1]{\left(#1\right)}
\newcommand{\brcs}[1]{\left\{#1\right\}}
\newcommand{\half}{\frac{1}{2}}

% Bold symbols and operators
\newcommand\Laplacian{\nabla^2}
\newcommand\bnabla{\boldsymbol{\nabla}}
\newcommand\bLaplacian{\boldsymbol{\nabla}^2}
\newcommand\bcdot{\boldsymbol{\cdot}}
\newcommand\bU{\mathscr{\boldsymbol{U}}}
\newcommand\bv{\boldsymbol{v}}
\newcommand\bV{\boldsymbol{V}}
\newcommand\bE{\boldsymbol{E}}
\newcommand\be{\boldsymbol{\hat{e}}}
\newcommand\bn{\boldsymbol{\hat{n}}}
\newcommand\bk{\boldsymbol{\hat{k}}}
\newcommand\bj{\boldsymbol{j}}
\newcommand\bi{\boldsymbol{i}}
\newcommand\bs{\boldsymbol{s}}
\newcommand\bA{\boldsymbol{A}}
\newcommand\bF{\boldsymbol{F}}
\newcommand\bI{\boldsymbol{I}}
\newcommand\bJ{\boldsymbol{J}}
\newcommand\bx{\boldsymbol{x}}
\newcommand\by{\boldsymbol{y}}
\newcommand\bz{\boldsymbol{z}}
\newcommand\br{\boldsymbol{r}}
\newcommand\bc{\boldsymbol{c}}
\newcommand\bxhat{\hat{\bx}}
\newcommand\byhat{\hat{\by}}
\newcommand\bzhat{\hat{\bz}}
\newcommand\brhat{\hat{\br}}
\newcommand\bnhat{\hat{\boldsymbol{n}}}
\newcommand\btheta{\boldsymbol{\theta}}
\newcommand\bthetahat{\hat{\btheta}}
\newcommand\bphi{\boldsymbol{\phi}}
\newcommand\bphihat{\hat{\bphi}}
\newcommand\bzero{\boldsymbol{0}}
\newcommand\bomega{\boldsymbol{\omega}}
\newcommand\bpsi{\boldsymbol{\psi}}

% Calligraphic symbols
\newcommand\cB{\mathcal{B}}
\newcommand\cE{\mathcal{E}}
\newcommand\cF{\mathcal{F}}
\newcommand\cO{\mathcal{O}}
\newcommand\cG{\mathcal{G}}
\newcommand\cI{\mathcal{I}}
\newcommand\cP{\mathcal{P}}
\newcommand\cD{\mathcal{D}}
\newcommand\cL{\mathcal{L}}
\newcommand\cU{\mathscr{U}}
\newcommand\cV{\mathscr{V}}
\newcommand\cW{\mathscr{W}}

% Tensors
\newcommand\tI{\mathsf{I}}
\newcommand\tS{\mathsf{S}}
\newcommand\tT{\mathsf{T}}

% Unit vector
\newcommand\ui{\boldsymbol{\hat{\imath}}}

\setlength{\textheight}{8.7in}
\setlength{\columnsep}{2.0pc}
\setlength{\textwidth}{6.6in}
% \setlength{\footheight}{0.2in}
\setlength{\topmargin}{0.05in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0.1in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
% \setlength{\parindent}{1pc}
\setlength{\parindent}{0.0 in}
\setlength{\parskip}{0.1 in}

\journal{Journal of Computational Physics}

\begin{document}

\begin{frontmatter}

\title{Computational Electrokinetics}
\author[tech]{Roman Zeyde}
\author[tech]{Irad Yavneh}

\address[tech]{Technion, Israel}

\begin{abstract}
This is the abstract.
\end{abstract}

\begin{keyword}
numerical \sep electrokinetics 
\end{keyword}

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations} \label{equations}
Electrokinetic theory describes the dynamics of charged particles
in ionic fluids \cite{kirby2010book}. When a particle acquires surface charge, a layer
of ions of opposite charge is attracted to the surface via
electric forces, creating a double-layer structure around the
particle (see Figure \ref{fig:EDL}). This structure, called
``Debye layer'', electrically screens the surface charge, thus
creating a potential difference between the particle and the outer
layer of the fluid bulk.
\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.25\textwidth]
            {figs/ElectricDoubleLayer.eps}
        \caption{Schematic structure of the double layer:
        (1) particle surface, (2) Debye layer and (3) fluid bulk.
        If the particle surface has positive (blue) surface charge,
        it attracts negative (red) ions from the fluid making the
        Debye layer negatively charged (as opposed to the rest of
        the fluid bulk, which is electrically neutral).
        The zeta potential is defined as the voltage drop across (2).}
        \label{fig:EDL}
    \end{center}
\end{framed}
\end{figure}
In cases where the layer width is much smaller than the particle's
size, an analytical asymptotic solution for the Debye layer
dynamics can be found \cite{yariv2010asymptotic}.

The partial differential equations that descibe the system dynamics
are coupled and non-linear. 
In general, they have no analytic solution. Moreover,
any numerical solver must handle the scale disparity caused by the
Debye layer width being much smaller than the particle's size. A
closed-form linear asymptotic solution has been developed for
spherical ion-exchanging particles and weak electric field in an
axisymmetric setting \cite{yariv2010migration} but it is hard to extend this
analytic solution to more general systems. Once the electric field
becomes stronger, significant nonlinear phenomena are expected.
This interesting regime has not yet been explored.

Recently, it has been conceived that such type of dynamics
can be very useful in transporting and manipulating micro-
and nanoscale objects for many nanotechnology applications,
such as the self-assembly of superstructures, roving sensors, 
drug-delivery systems and useful nanomachinery 
\cite{howse2007self,paxton2004catalytic,pumera2010electrochemically}.
However, since the analytic model for these electrokinetic
phenomena has no closed-form analytic solution, a numerical
solver has to be developed for analysing such system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Micro-scale phenomena}

\subsubsection{Dimensionless Notation}

Spatial coordinates $\br$ are normalized by $a^*$ (the characteristic size of the ion-exchanger).

Positive and negative ionic concentrations are denoted by $c_+$ and $c_-$ respectively, and
are normalized by ambient ionic concentration $c^*$. The ionic valences are $\pm z$.

The dimensional ionic fluxes are given by Nernst-Plank equation 
(using Einstein relation for the electric mobility, $\mu_q^* = D^* / k_B^* T^*$):
\begin{eqnarray}
\bj^*_\pm &=& 
-D^* \bnabla c^*_\pm + \bv^* c^*_\pm \mp \frac{z e^* D^*}{k_B^* T^*} c^*_\pm \bnabla \varphi^*,
\end{eqnarray}
where $D^*$ is the ionic diffusivity.

The electric potential $\varphi$ is normalized by the thermal voltage:
\begin{eqnarray}
\varphi^* &=& \frac{k_B^* T^*}{z e^*} = \frac{R^* T^*}{z F^*},
\end{eqnarray}
where $R^* = k_B^* N_A$ and $F^* = e^* N_A$ is the Faraday number.

The dimensional Stokes equation:
\begin{eqnarray}
-\bnabla p^* + \mu^* \bLaplacian \bv^* + \eps^* \Laplacian \varphi^* \bnabla \varphi^* &=& 0.
\end{eqnarray}

The velocity $\bv$ is normalized by:
\begin{eqnarray}
v^* &=& \frac{\eps^* (\varphi^*)^2}{a^* \mu^*},
\end{eqnarray}
and the pressure $p$ is normalized by:
\begin{eqnarray}
p^* &=& \frac{\mu^* v^*}{a^*} = \frac{\eps^* (\varphi^*)^2}{(a^*)^2}.
\end{eqnarray}

The fluxes $\bj^*_\pm$ are normalized by:
\begin{eqnarray}
j^* = \frac{D^* c^*}{a^*}.
\end{eqnarray}

The nondimensional Nernst-Plank equation is written as:
\begin{eqnarray}
\bj_\pm &=& 
-\bnabla c_\pm + \alpha \bv c_\pm \mp c_\pm \bnabla \varphi, \\
\alpha &=&
\frac{v^* a^*}{D^*} = \frac{\eps^* (\varphi^*)^2}{\mu^* D^*}.
\end{eqnarray}
where $\alpha$ is the dimensionless Peclet number.

The dimensional Poisson equation is:
\begin{eqnarray}
(c^*_+ - c^*_-) z F^* &=& -\eps^* \Laplacian \varphi^*,
\end{eqnarray}
and after nondimensionalization, is:
\begin{eqnarray}
c_+ - c_- &=& -\frac{\eps^* \varphi^*}{z F^* c^* (a^*)^2} \Laplacian \varphi = 
-\delta^2 \Laplacian \varphi,
\\
\delta^2 &=& \frac{\eps^* \varphi^*}{c^* F^* z (a^*)^2} = 
\pars{\frac{\delta^*}{a^*}}^2.
\end{eqnarray}

The force is denoted by $\bF$ normalized by $\mu^* v^* a^* = \eps^* (\varphi^*)^2$.
The stress tensor is denoted by $\tT$ and normalized by $p^* = \mu^* v^* / a^*$.
The electric field is denoted by $\bE$ and normalized by $E^* = \varphi^* / a^*$.

\subsubsection{Nernst-Planck Equations}
The ionic fluxes are given by ion diffusion, electrostatic forces and ion advection by the fluid:
\begin{eqnarray}
  \bj_\pm &=& -\bnabla c_\pm \mp c_\pm \bnabla \varphi + \alpha \bv c_\pm,
\end{eqnarray}

Due to conservation of ions, the fluxes are divergence-free:
\begin{eqnarray}
\bnabla \cdot \bj_\pm &=& 0.
\end{eqnarray}
These equation may be re-written by using the following transform:
\begin{eqnarray}
  c &=& \frac{c_+ + c_-}{2},\\
  q &=& \frac{c_+ - c_-}{2}.
\end{eqnarray}
Thus, salt and charge fluxes are defined by:
\begin{eqnarray}
  \bj &=& \frac{\bj_+ + \bj_-}{2} = -\bnabla c - q \bnabla \varphi + \alpha \bv c, \\
  \bi &=& \frac{\bj_+ - \bj_-}{2} = -\bnabla q - c \bnabla \varphi + \alpha \bv q,
\end{eqnarray}
and are divergence-free as well:
\begin{eqnarray}
\bnabla \cdot \bj &=& 0, \\
\bnabla \cdot \bi &=& 0. 
\end{eqnarray}

\subsubsection{Incompressible Stokes Flow with Electrostatic Force}
Mass conservation (due to constant fluid density):
\begin{eqnarray}
\bnabla \cdot \bv &=& 0.
\end{eqnarray}
Momentum conservation (force balance):
\begin{eqnarray}
\Laplacian \bv - \bnabla p + \bnabla \varphi \Laplacian \varphi &=& 0.
\end{eqnarray}
Free charge density is given by the Poisson equation:
\begin{eqnarray}
\delta^2 \Laplacian \varphi &=& -2q.
\end{eqnarray}

\subsubsection{Boundary conditions}
An constant electric field $\bE$ is applied to the system, such that $|\bE| = \beta$.
The ion-exchanger will acquire a non-zero steady-state velocity $\cU$, which is
a function of the applied field, and is not known a priori.

Far away from the particle, we have uniform applied electrostatic field, uniform flow 
and ambient ionic concentrations:
\begin{eqnarray}
\bv &\rightarrow& -\cU \ui, \\
\bE = -\bnabla \varphi &\rightarrow& \beta\ui, \\
c_\pm &\rightarrow& 1.
\end{eqnarray}

We consider a spherical ion-exchanger. 
At its surface $\mathcal{S}$ (defined at $r=1$ with normal $\bn = \brhat$), 
we have zero slip, anion impermeability, cation selectivity (with fast Butler-Volmer kinetics)
and high conductance of the particle (yielding uniform electric potential):
\begin{eqnarray}
\bv & = & \bzero, \\
\bn \cdot \bj_- & = & 0, \\
\bn \cdot \bj_+ & = & k (1 - c_+/\gamma) \Rightarrow c_+ \rightarrow \gamma, \\
\varphi & = & \cV.
\end{eqnarray}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Separation of scales}
When $\delta \ll 1$, 
most of the fluid remains electroneutral ($q \approx 0$, thus $c_+ \approx c_-$), 
except a thin boundary layer of width $O(\delta)$, 
also known as the ``Debye layer'', surrounding the ion exchanger.

\subsubsection  {Bulk-scale equations}
The Nernst-Planck equations above for the fluid bulk (outside
the boundary layer), using bulk variables, read:
\begin{eqnarray}
  Q & = & 0, \\
  C & = & C_+ = C_-, \\
\bJ &=& -\bnabla C + \alpha \bV C, \\
\bI &=& -C \bnabla \varPhi.
\end{eqnarray}
Because the flow is incompressible ($\bnabla \cdot V = 0$), 
ion conservation results in the following equations:

\begin{enumerate}
\item Salt conservation macroscale equation:
\begin{eqnarray}
\bnabla \cdot \bJ = \Laplacian C - \alpha \bV \cdot \bnabla C = 0. 
\end{eqnarray}

\item Charge conservation macroscale equation:
\begin{eqnarray}
\bnabla \cdot \bI = \bnabla \cdot \pars{ C \bnabla \varPhi } = 0.
\end{eqnarray}
\end{enumerate}

The macroscale Stokes flow equations have the same form as the microscale equations:
\begin{eqnarray}
\bnabla \cdot \bV &=& 0, \\
\Laplacian \bV - \bnabla P + \bnabla \varPhi \Laplacian \varPhi &=& 0.
\end{eqnarray}

\subsubsection{Debye-scale effective boundary conditions}
We define the scaled coordinate $\rho$ (normal to ion-exchanger surface), where $\delta \ll 1$:
\begin{eqnarray}
  \rho &=& \frac{r-1}{\delta}. 
\end{eqnarray}
Thus, the electric potential and the ionic concentrations are $O(1)$ at the Debye layer ($r \approx 1$):
\begin{eqnarray}
  \varphi(\rho,\theta) &\sim& \varphi, \\
  c_\pm(\rho,\theta) &\sim& c_\pm, \\
  c(\rho,\theta) &\sim& c, \\
  q(\rho,\theta) &\sim& q.
\end{eqnarray}

The radial fluxes at the Debye layer are $O(\delta^{-1})$, due to coordinate rescaling:
\begin{eqnarray}
  j_\pm(\rho, \theta) &\sim& \delta^{-1} j_\pm.
\end{eqnarray}

The pressure is $O(\delta^{-2})$, due to momentum balance equations:
\begin{eqnarray}
  p(\rho, \theta) &\sim& \delta^{-2} p.
\end{eqnarray}

The tangential velocity component is $O(1)$ and the radial component is $O(\delta)$.

In order to match bulk $O(1)$ outer radial ionic flux, 
the $O(\delta^{-1})$ term of the inner ionic radial flux, $j_\pm(\rho, \theta)$ must vanish:
\begin{eqnarray}
  -\deriv{c_\pm}{\rho} \mp c_\pm \deriv{\varphi}{\rho} &=& 0, \\
  \deriv{\varphi}{\rho} \pm\frac{1}{c_\pm}\deriv{c_\pm}{\rho} &=& 0, \\
  \pars{\varPhi - \varphi} \pm\pars{\log C_\pm - \log c_\pm} &=& 0, \\
  \varPhi - \varphi \pm \log \frac{C_\pm}{c_\pm} &=& 0.
\end{eqnarray}
This relation can be rewritten as a Boltzmann distribution of ionic concentration:
\begin{eqnarray}
\lim_{\rho\rightarrow\infty} c_\pm &=& C, \\
c_\pm &=& C \exp\left[\pm(\varPhi - \varphi)\right].
\end{eqnarray}

Due to cation selectivity, $c_+|_{\rho=0} = \gamma$.
\begin{eqnarray}
  c_+ = \gamma &=& C \exp(\varPhi - \cV), \\
  \cV + \log \gamma &=& \log C + \varPhi.
\end{eqnarray}

The so-called zeta potential is 
the voltage drop between the particle surface and the end of
the boundary layer, denoted $\zeta = \cV - \varPhi$.

Without loss of generality, 
we assume that $\cV = -\log\gamma$, thus $\zeta = -\log\gamma-\varPhi$:
\begin{eqnarray}
\varPhi + \log C &=& 0.
\end{eqnarray}

Due to anion impermeability and radial flux continuity:
\begin{eqnarray}
  \br \cdot \bj_- = -\deriv{c_-}{r} + c_- \deriv{\varPhi}{r} &=& 0, \\
-\deriv{C}{r} + C \deriv{\varPhi}{r} &=& 0, \\
\deriv{}{r}\pars{\varPhi - \log C} &=& 0.
\end{eqnarray}

Using the derivation in Appendix \ref{append:slip} and the fact that $V_r = 0$ 
on the boundary, the slip condition can be written using the surface
gradient operator:
\begin{eqnarray}
\bV &=& 
\zeta \cdot \bnabla_\mathcal{S} \varPhi 
+ 2\log\pars{1-\tanh^2\frac{\zeta}{4}} \cdot \bnabla_\mathcal{S} \log C, \\
\bnabla_\mathcal{S} &=& \pars{\tI - \bn\bn} \bnabla.
\end{eqnarray}

\subsection{Steady state}
The stress tensor is composed of Newtonian and Maxwel stresses:
\begin{eqnarray}
\tT &=& \bnabla \bV + (\bnabla \bV)^\dagger - P \tI
+ \bnabla \varPhi \bnabla \varPhi - \frac{1}{2} (\bnabla \varPhi \cdot \bnabla \varPhi) \tI, \\
\bnabla \cdot \tT &=& \bzero.
\end{eqnarray}
The total force acting on the particle is computed by the following surface integral:
\begin{eqnarray}
\bF &=& \oint_\mathcal{S} \tT \cdot \bnhat dA.
\end{eqnarray}
In steady state (constant particle drift velocity $\cU$), 
the total force must be zero: 
\begin{eqnarray}
\bF = \bzero.
\end{eqnarray}

The total force $\bF(\beta, \cU)$ is a function of the applied electric field $\beta$ and
the velocity $\cU$.
For any given electric field $\beta$, the force-free constraint above 
can be solved numerically for $\cU$ to
yield the corresponding steady-state velocity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algorithm Description} \label{algorithm}
\subsection{Discretization}
\subsubsection{Spherical Coordinates}
Because the system is axisymmetric, the operators are written in 
spherical coordinates $(r,\theta,\phi)$ 
(using the derivation of Appendix \ref{append:spherical}).

Due to axial symmetry, 
the scalar gradient and the divergence can be written as:
\begin{eqnarray}
\bnabla f &=& \deriv{f}{r}\brhat + \frac{1}{r}\deriv{f}{\theta}\bthetahat, \\
\bnabla \cdot \bF &=& \frac{1}{r^2}\deriv{}{r}\pars{F_r r^2 } + 
               \frac{1}{r \sin\theta}\deriv{}{\theta}\pars{F_\theta \sin\theta}.
\end{eqnarray}
The scalar Laplacian can be written as:
\begin{eqnarray}
\Laplacian f = \bnabla \cdot (\bnabla f)&=& 
 \frac{1}{r^2}\deriv{}{r}\pars{r^2 \deriv{f}{r}} + 
 \frac{1}{r^2 \sin\theta}\deriv{}{\theta}\pars{\sin\theta \deriv{f}{\theta}}.
\end{eqnarray}
The vector gradient can be written as:
\begin{eqnarray}
\bnabla \bF &=& \deriv{F_r}{r} \brhat \brhat + \deriv{F_\theta}{r} \brhat \bthetahat + 
\frac{1}{r}\pars{\deriv{F_r}{\theta} - F_\theta} \bthetahat \brhat + 
\frac{1}{r}\pars{\deriv{F_\theta}{\theta} + F_r} \bthetahat \bthetahat.
\end{eqnarray}
The vector Laplacian can be written as:
\begin{eqnarray}
\bLaplacian \bF &=& 
\left(\Laplacian F_r - \frac{2F_r}{r^2} - 
\frac{2}{r^2 \sin\theta} \deriv{\left(F_\theta \sin\theta \right)}{\theta}\right)\brhat
+ \left(\Laplacian F_\theta - \frac{F_\theta}{r^2 \sin^2\theta} + 
\frac{2}{r^2}\deriv{F_r}{\theta}\right) \bthetahat.
\end{eqnarray}

\subsubsection{Computational Grid}
In order to discretize the differential operators, 
a regular grid of size $n_r \times n_\theta$ is defined, with:
\begin{eqnarray}
(r_i,\theta_j) &\in& [1, \infty) \times [0,\pi], \\ 
r_i &=& (1+\Delta_r)^i, \\
\theta_j &=& \Delta_\theta \cdot j,
\end{eqnarray}
where a logarithmic grid spacing is used for $r$ and a uniform grid is used for $\theta$.
Note that $r_0 = 1$ and $R_{max} = (1+\Delta_r)^{n_r}$ should be much 
larger than $r_0$. Thus:
\begin{eqnarray}
\Delta_r &=& \pars{R_{max}} ^ \frac{1}{n_r} - 1, \\
\Delta_\theta &=& \frac{\pi}{n_\theta}.
\end{eqnarray}

This grid induces a disjoint subdivision of the domain 
$\Omega = [1, R_{max}] \times [0,\pi]$ into cells:
\begin{eqnarray}
\bigcup_{ij}\Omega_{ij} &=& \Omega.
\end{eqnarray}
A specific cell $\Omega_{ij}$ and its center $(\bar{r}_i, \bar{\theta}_j)$ are defined by:
\begin{eqnarray}
\Omega_{ij} &=& [r_{i-1}, r_{i}] \times [\theta_{j-1}, \theta_{j}], \\
\bar{r}_i &=& \frac{r_{i-1} + r_{i}}{2}, \\
\bar{\theta}_j &=& \frac{\theta_{j-1} + \theta_{j}}{2}.
\end{eqnarray}

Each variable is discretized with respect of its cell as follows:
\begin{itemize}
\item $\varPhi$, $C$ and $P$ are represented by their value at the center of each cell, 
using an all-centered grid:
\begin{eqnarray}
\varPhi^h[i,j] &=& \varPhi(\bar{r}_i, \bar{\theta}_j), \\
C^h[i,j] &=& C(\bar{r}_i, \bar{\theta}_j), \\
P^h[i,j] &=& P(\bar{r}_i, \bar{\theta}_j).
\end{eqnarray}
\item $\bV$ is represented by its values at cell boundaries, using a staggered grid:
\begin{eqnarray}
V_r^h[i,j] &=& V_r(r_i, \bar{\theta}_j), \\
V_\theta^h[i,j] &=& V_\theta(\bar{r}_i, {\theta}_j).
\end{eqnarray}
\end{itemize}

\subsubsection{Operator Discretization}
A finite-volume method with linear interpolation is used for flux 
$\boldsymbol{f}^h$ discretization.

Define the following discrete central difference operators:
\begin{eqnarray}
\cD_r(f^h)[i,j] &=& \frac{f^h\left[i+\half,j\right] - f^h\left[i-\half,j\right]}
                       {r\left[i+\half,j\right] - r\left[i-\half,j\right]}, \\
\cD_\theta(f^h)[i,j] &=& \frac{f^h\left[i,j+\half\right] - f^h\left[i,j-\half\right]}
					   {\theta\left[i,j+\half\right] - \theta\left[i,j-\half\right]}.
\end{eqnarray}

Define the following interpolation operators:
\begin{eqnarray}
\cI_r(f^h)[i,j] &=& \frac{
\pars{r\left[i,j\right] - r\left[i-\half,j\right]} 
  f^h\left[i+\half,j\right] + 
\pars{r\left[i+\half,j\right] - r\left[i,j\right]} 
  f^h\left[i-\half,j\right] 
}{r\left[i+\half,j\right] - r\left[i-\half,j\right]},
\\
\cI_\theta(f^h)[i,j] &=& 
\frac{
\pars{r\left[i,j\right] - r\left[i,j-\half\right]} 
  f^h\left[i,j+\half\right] + 
\pars{r\left[i,j+\half\right] - r\left[i,j\right]} 
  f^h\left[i,j-\half\right] 
}{r\left[i,j+\half\right] - r\left[i,j-\half\right]}.
\end{eqnarray}

The total flux of a cell is equal to zero (due to conservation):
\begin{eqnarray}
\bnabla \cdot \boldsymbol{f}^h &=& 0, 
\\
\frac{1}{r^2} \cD_r\pars{f^h_r r^2} + 
\frac{1}{r \sin\theta} \cD_\theta\pars{f^h_\theta \sin\theta} &=& 0, 
\\
\cD_r\pars{f^h_r \cdot r^2 \sin\theta} + 
\cD_\theta\pars{f^h_\theta \cdot r \sin\theta} &=& 0. 
\end{eqnarray}

Ion fluxes are discretized on grid cell boundaries:
\begin{eqnarray}
I^h_r &=& -\cI_r(C^h) \cdot \cD_r(\varPhi^h), \\
I^h_\theta &=& -\cI_\theta(C^h) \cdot \frac{\cD_\theta(\varPhi^h)}{r}.
\end{eqnarray}

Salt fluxes are discretized on grid cell boundaries 
(using upwind scheme $\mathcal{U}$ for numerical stability at large cell Peclet number):
\begin{eqnarray}
J^h_r &=& -\cD_r(C^h) + \alpha V^h_r \cdot \mathcal{U}^{\bV^h}_r (C^h), \\
J^h_\theta &=& -\frac{\cD_\theta(C^h)}{r} + \alpha V^h_\theta \cdot \mathcal{U}^{\bV^h}_\theta (C^h), \\
 \mathcal{U}^{\bV^h}_r(C^h)[i,j] &=& C^h\left[i-\frac{\sign(V^h_r)}{2}, j\right], \\
 \mathcal{U}^{\bV^h}_\theta(C^h)[i,j] &=& C^h\left[i, j-\frac{\sign(V^h_\theta)}{2}\right]. 
\end{eqnarray}

Mass flux $\bV$ is discretized on grid cells boundaries, 
using the velocity staggered grid.

Force components are discretized on the velocity staggered grid, where
linear interpolation is used for the Coulomb force.
\begin{eqnarray}
F^h_r &=& -\cD_r(P^h) 
          + \cL(V^h_r) - \frac{2}{r^2} V^h_r 
		  - \frac{2}{r^2 \sin\theta} \cI_r(\cD_\theta (V^h_\theta \sin\theta))
          + \cD_r(\varPhi^h) \cdot \cI_r(\cL(\varPhi^h)), \\
F^h_\theta &=& -\frac{\cD_\theta(P^h)}{r} 
		  + \cL(V^h_\theta) - \frac{F^h_\theta}{r^2 \sin^2\theta} 
		  + \frac{2}{r^2} \cI_\theta(\cD_\theta(F^h_r))
		  + \frac{\cD_\theta(\varPhi^h)}{r} \cdot \cI_\theta(\cL(\varPhi^h)), \\
\cL(f^h) &=& \frac{1}{r^2}\cD_r\pars{\cD_r(f^h) r^2} + 
\frac{1}{r^2 \sin\theta} \cD_\theta\pars{\cD_\theta(f^h) \cdot \sin\theta}.
\end{eqnarray}

\subsubsection{Boundary Conditions}
In order to discretize boundary condtions, 
``ghost'' points are employed. 
The grid is extended to include points outside the domain interior,
and the variables at the ``ghost'' points are set to satisfy 
the discretized boundary conditions, following the analysis below.

The boundary conditions at the ion-exchange boundary ($r=1$) are:
\begin{eqnarray}
\cI_r(\varPhi^h) &=& -\cI_r(\log C^h), \\
\cD_r(\varPhi^h) &=& \cD_r(\log C^h), \\
V^h_r &=& 0, \\
V^h_\theta &=& 4\log\pars{\frac{1 + \exp\left\{\cI_\theta(\zeta^h)/2\right\}}{2}} \cdot 
			\cD_\theta(\zeta^h), \\
  \zeta^h &=& - \log \gamma - \cI_r(\varPhi^h).
\end{eqnarray}
The equations are discretized using ``ghost'' points outside the interior of the domain:
\begin{eqnarray}
 \varPhi^h[1/2, j] = \frac{\varPhi^h[0,j] + \varPhi^h[1,j]}{2} &=& 
	-\frac{\log C^h[0,j] + \log C^h[1,j]}{2} = -\log C^h[1/2, j], \\
  \frac{\varPhi^h[1,j] - \varPhi^h[0,j]}{\Delta r} &=& 
	\frac{\log C^h[1,j] - \log C^h[0,j]}{\Delta r}. 
\end{eqnarray}
By adding and subtracting the equations, we have:
\begin{eqnarray}
\varPhi^h[0,j] &=& -\log\brcs{C^h[1,j]}, \\
C^h[0,j] &=& \exp\brcs{-\varPhi^h[1,j]}, \\
  \zeta^h[j] &=& \cV - \varPhi^h[1/2,j]
                     = -\log\gamma -\frac{\varPhi^h[0,j] + \varPhi^h[1,j]}{2}.
\end{eqnarray}
The slip condition can be written as:
\begin{eqnarray}
V_r^h[0,j] &=& 0, \\
V_\theta^h[1/2,j] &=& 
4\log\pars{\frac{1 + \exp\left\{\frac{1}{2}
 \frac{\zeta^h[j] + \zeta^h[j+1]}{2}\right\}}{2}} 
\cdot \frac{\zeta^h[j+1] - \zeta^h[j]}{\bar{\theta}_{j+1} - \bar{\theta}_{j}}, \\
  V^h_\theta[0,j] &=& 2 V^h_\theta[1/2, j] - V^h_\theta[1,j].
\end{eqnarray}

For $\theta = 0$ and $\theta = \pi$, the ``ghost'' points are defined by:
\begin{eqnarray} 
\brc{l}{
\varPhi^h[i, 0] = \varPhi^h[i, 1] \\
C^h[i, 0] = C^h[i, 1] \\
V_r^h[i, 0] = V_r^h[i, 1] \\
V_\theta^h[i, 0] = 0
} 
\brc{l}{
\varPhi^h[i, n_\theta+1] = \varPhi^h[i, n_\theta], \\
C^h[i, n_\theta+1] = C^h[i, n_\theta], \\
V_r^h[i, n_\theta+1] = V_r^h[i, n_\theta], \\
V_\theta^h[i, n_\theta+1] = 0.
} 
\end{eqnarray}

Far away from the ion-exchange boundary ($r\rightarrow\infty$) we have:
\begin{eqnarray}
\frac{\varPhi^h[n_r + 1, j] - \varPhi^h[n_r, j]}{\bar{r}_{n_r + 1} - \bar{r}_{n_r}}, 
 & = & -\beta \cos\bar{\theta}_j, \\
C^h[n_r + 1, j] & = & 1 \\
V_r^h[n_r + 1, j] & = & -\cU \cos\bar{\theta}_j, \\
V_\theta^h[n_r + 1, j] & = & \cU \sin\theta_j.
\end{eqnarray}

Because the actual grid is finite, the choice 
$r_{n_r} = R_{max}$ must be taken large enough so as to have a negligible effect
on the solution.

Note that $P^h$ has no boundary conditions -- so the pressure variable is defined 
only in the domain interior.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solver Design}
\subsubsection{Newton's Method}
An operator $\cO$ is defined by its input vector $\bx$ and its output 
$\by = \cO(\bx)$.

In order to solve the equation $\cO(\bx) = \bzero$, Newton's method can 
be applied, given an initial solution $\bx_0$:
\begin{eqnarray}
\bzero = \cO(\bx_n + \Delta \bx_n) &\approx& \cO(\bx_n) + \bnabla \cO(\bx_n)\cdot\Delta\bx_n, \\
\Rightarrow \bnabla \cO(\bx_n) \cdot \Delta \bx_n &=& -\cO(\bx_n), \\
\bx_{n+1} &=& \bx_{n} + \Delta \bx_n. 
\end{eqnarray}

Each step requires the computation of the residual vector, $\cO(\bx)$, and
the gradient matrix of the operator, $\bnabla \cO(\bx)$, 
given the current solution $\bx = \bx_n$.
Then, a sparse linear system needs to be solved to yield the update for the 
increment $\Delta \bx_n = \bx_{n+1} - \bx_n$.

When the method converges, $\bx_n \rightarrow \bx_\infty$, 
the convergence is in general quadratic:
\begin{eqnarray}
\|\bx_{n+1} - \bx_\infty\| \le k \|\bx_{n} - \bx_\infty\|^2.
\end{eqnarray}
Thus, when the initial guess is close enough to the solution, the solver
will typically converge in very few steps.

\subsubsection{Operator Representation}
System variable $\bx$ is defined as the concatenation $[\,\varPhi, C, V_r, V_\theta, P\,]$,
taking only values in the interior of the problem domain grid.
Thus, each problem variable can be computed by applying an appropriate projection operator on $\bx$:
\begin{eqnarray}
X &=& \cP_X(\bx) \mbox{ for } X \in \left\{\varPhi, C, V_r, V_\theta, P\right\}.
\end{eqnarray}

The extended variables $\tilde X$ (including ``ghost'' point values) 
are computed by applying the following nonlinear operators:
\begin{eqnarray}
\tilde{\varPhi} &=& \cB_\varPhi(\varPhi, C; \beta), \\
\tilde{C} &=& \cB_C(C, \varPhi), \\
\tilde{V_r} &=& \cB_{V_r}(V_r; \cU), \\
\tilde{V_\theta} &=& \cB_{V_\theta}(V_\theta, C, \varPhi; \cU, \gamma), \\
\tilde{P} &=& P.
\end{eqnarray}
The extended system variable can be written as $\tilde{\bx} = \cB(\bx)$, 
where $\cB$ is the ``ghost'' point extension operator (derived using boundary conditions)
so the concatenated equations of the system (as described above) can be written 
as a non-linear operator $\cE(\tilde\bx) = \bzero$ acting on the extended system variable 
$\tilde{\bx}$, where $\cE$ is the conservation equations operator 
(derived by concatenation of the system equations).

The problem can be cast as the following non-linear system:
\begin{eqnarray}
\bzero = \cE (\tilde{\bx}) = \cE (\cB (\bx)) = \cO(\bx),
\end{eqnarray}
 where $\cO = \cE \circ \cB$.

\subsubsection{Steady-state}
Assume that $\cU \ui$ is the velocity of the ion exchanger 
(given that the fluid is at rest far away from the particle).
After the convergence of the solver, $\cO(\bx_n) \rightarrow \bzero$, 
the ``ghost'' point values are updated using $\tilde\bx = \cB(\bx_\infty)$,
so  that the surface force $\boldsymbol{f} = \tT \cdot \bn$ acting 
on the ion-exchanger surface
can be computed by integration over the ion exchanger's surface $\mathcal S$.

Due to symmetry considerations, the total force $\bF$ is 
aligned with $\ui$ and it vanishes iff $\ui \cdot \bF = 0$.
\begin{eqnarray}
\bF &=& \oint_\mathcal{S} \boldsymbol{f}  dA = 
\oint_\mathcal{S} \tT \cdot \bnhat  dA. \\
F_\imath = \ui \cdot \bF &=& \oint_\mathcal{S} \ui \cdot \boldsymbol{f}  dA = 
\oint_\mathcal{S} \ui \cdot \tT \cdot \bnhat  dA = 
\int_0^\pi f_\imath(\theta) \cdot 2\pi \sin\theta d\theta ,
\\  
f_\imath &=& \ui \cdot \tT \cdot \brhat,
\\  
f_\imath &=& \pars{-P + 2\cD_r(V_r) + 
\frac{1}{2}\pars{\cD_r(\varPhi)}^2 - \frac{1}{2r^2}\pars{\cD_\theta(\varPhi)}^2}\cos\theta 
\\  
&& -\pars{\cD_r(V_\theta) - \frac{V_\theta}{r}
+ \frac{1}{r}\cD_r(\varPhi) \cD_{\theta}(\varPhi)}\sin\theta, \\
 \mathcal{S} &=& \{\br : \|\br\|_2 = 1\}.
\end{eqnarray}

The integral for $F_\imath$ is approximated by 1D numerical quadrature
(using the mid-point rule), to yield 
the total force $F_\imath(\cU)$ as a function of the drift velocity $\cU$.

Since the goal is to find steady-state solution, $F_\imath(\cU)$ 
is required to be zero --
and the appropriate $\cU$ is found by a simple 1D root-finding algorithm,
applied as an outer loop.

\subsubsection{Continuation}

The iterative solver above can be used to compute the steady-state solution for any given $\beta$.
For $\beta \ll 1$, the linear terms are the dominant ones, so the solution is linear in $\beta$ 
(as derived in \cite{yariv2010migration}), and the solver convergence is fast.

However, it is no longer true for $\beta \sim O(1)$, since the nonlinear terms become dominant
and the iterative solver may not converge at all.
In order to find a solution for such $\beta$, continuation method is used:
the solver is applied to a sequence of $\{\beta_i\}_{i=0}^n$ such that $\beta_0 = 0$,
$\beta_n = \beta$ and the solution $\bx_i$ for $\beta_i$ is used as the solver initializer
for the problem of $\beta_{i+1}$.

Given that $|\beta_{i+1} - \beta_i| < \eps$ for small enough $\eps$, 
the next solution $\bx_{i+1}$ is close to the previous one $\bx_i$,
so the iterative solver converges well.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solver Implementation}

Object-Oriented Design methodology is used for implementation.
The MATLAB programming language is chosen due to strong numerical capabilities
and high-level language features.

\subsubsection{Operator Interface $ \rightarrow \cO$}
The base \verb|Operator| interface is defined by 
supporting the computation of a residual vector \verb|op.res()|
and a gradient matrix \verb|op.grad()|, given an input vector $\bx$.

This interface is implemented by specific operators classes,
which are used to constuct the operator $\cO = \cE \circ \cB$, so that 
the Newton method solver (that requires computing the residual and the gradient) 
can be applied automatically.

Each such operator may have other operators as its inputs, so the residual
and the gradient are computed recursively, using the chain rule.

\paragraph{Linear operator}
If $\cO$ is linear, it can be represented by a matrix $L$, such that:
\begin{eqnarray}
\cO(\bx) &=& L \bx, \\
\bnabla\cO &=& L. 
\end{eqnarray}
Note that finite difference $\cD$, interpolation $\cI$ and 
upwind selection $\mathcal{U}$ operators
can be implemented as linear sparse operators, by constructing 
an appropriate sparse matrix $L$, having $O(\dim \bx)$ non-zeroes).

\paragraph{Pointwise scalar function}
Let $f: \R \rightarrow \R$ be a differentiable 1D function, whose derivative is
denoted by $f': \R \rightarrow \R$. The operator $\cF$ can be defined to
represent pointwise application of $f$:
\begin{eqnarray}
\cF(\bx) &=& [f(x_1); \ldots; f(x_n)], \\
\bnabla\cF(\bx) &=& \diag\{f'(x_1), \ldots, f'(x_n)\}.
\end{eqnarray}
$f$ is defined and differentiated automatically using the MATLAB symbolic toolbox.

\paragraph{Constant value}
A constant operator $\mathcal{C}$ is defined as having constant vector value $\bc$ 
(that does not depend on $\bx$):
\begin{eqnarray}
\mathcal{C}(\bx) &=& \bc, \\
\bnabla\mathcal{C} &=& \bzero.
\end{eqnarray}

\paragraph{Binary operators}
Let $\cO_1$ and $\cO_2$ be two operators, so their pointwise addition, subtraction,
and multiplication, are defined as follows:
\begin{eqnarray}
(\cO_1 \pm \cO_2)(\bx) &=& \cO_1)(\bx) \pm \cO_2(\bx), \\
\bnabla (\cO_1 \pm \cO_2)(\bx)   &=& \bnabla\cO_1(\bx) \pm \bnabla\cO_2(\bx), \\
(\cO_1 \cdot \cO_2)(\bx) &=& \diag(\diag(\cO_1(\bx)) \cdot \diag(\cO_2(\bx))), \\
\bnabla (\cO_1 \cdot \cO_2)(\bx) &=& \diag(\cO_2(\bx)) \bnabla\cO_1(\bx) 
                                  + \diag(\cO_1(\bx)) \bnabla\cO_2(\bx).
\end{eqnarray}

\paragraph{N-ary operators}
Given $N$ operators $\{\cO_i\}_{i=1}^N$, their concatenation $\cO$ is defined as:
\begin{eqnarray}
\cO(\bx) &=& [\cO_1(\bx); \ldots; \cO_N(\bx)], \\
(\bnabla\cO)(\bx) &=& [(\bnabla\cO_1)(\bx); \ldots; (\bnabla\cO_N)(\bx)].
\end{eqnarray}

\subsubsection{Iterative Newton solver $\rightarrow \bx$}
Given an initial solution $\bx_0$ and specific values for $\beta$ and $\cU$, 
we apply the following algorithm 
until convergence ($\|\Delta \bx\| < \eps$):
\begin{enumerate}
\item Update the system: $\verb|op| = \verb|System|(\bx_{n})$
\item Compute the gradient: $\boldsymbol{A}_n = \verb|op.grad()|$
\item Compute the right-hand side: $\boldsymbol{b}_n = \verb|-op.res()|$
\item Solve sparse linear system: $\boldsymbol{A}_n \Delta \bx_n = \boldsymbol{b}_n$
\item Update solution: $\bx_{n+1} = \bx_{n} + \Delta \bx_{n}$
\end{enumerate}

\subsubsection{Total force integration $\rightarrow F_\imath$}
After the iterative Newton solver converges ($\bx_n \rightarrow \bs$), 
the ``ghost'' point values are computed by computing $\cB(\bs)$ and
the total force $\bF$ acting on the ion-exchanger.

$F_\imath = \bF \cdot \ui$ is computed by using the mid-point quadrature rule:
\begin{eqnarray}
F_\imath &=& \sum_{i=1}^{n_\theta} f(\bar\theta_i) \cdot 
              2 \pi \sin\bar\theta_i \cdot \Delta\theta_i
\end{eqnarray}

\subsubsection{Steady state drift velocity $\rightarrow \cU$}
In steady state, the total force acting on the particle $F_\imath$ must vanish.
This constraint yields an equation, whose solution is the steady-state velocity $\cU$:
\begin{eqnarray}
F_\imath(\cU) &=& 0
\end{eqnarray}
This equation can be solved using a 1D root finding algorithm, e.g. the secant method:
\begin{eqnarray}
\cU_{n+1} = \frac{F_\imath(\cU_{n})\cU_{n-1} - F_\imath(\cU_{n-1})\cU_{n}}
{F_\imath(\cU_{n}) - F_\imath(\cU_{n-1})} \rightarrow \cU
\end{eqnarray}
The result $\cU$ is the 
steady state particle drift velocity for the electric field $\beta$.

The non-linear system is defined by the following PDEs:
\label{PDEs}
\begin{eqnarray}
\bnabla \cdot \pars{C \bnabla \varPhi} &=& 0, \\
\bnabla \cdot \pars{\bnabla C - \alpha \bV C} &=& 0, \\
\bLaplacian \bV - \bnabla P + \Laplacian \varPhi \bnabla \varPhi &=& \bzero.
\end{eqnarray}

For $\zeta(\theta) = \cV - \varPhi(1, \theta)$ and 
$\cV = -\log \gamma$, the boundary conditions on $r = 1$ are:
\begin{eqnarray}
\varPhi + \log C &=& \cV + \log \gamma = 0, \\
\deriv{}{r}\pars{\varPhi - \log C} &=& \frac{J_-}{C} = 0, \\
\bV = \zeta \cdot \bnabla_\mathcal{S} \varPhi 
+ 2\log\pars{1-\tanh^2\frac{\zeta}{4}} \cdot \bnabla_\mathcal{S} \log C
&=& 4\log\pars{\frac{1 + e^\frac{\zeta}{2}}{2}} \bnabla_S \varPhi.
\end{eqnarray}

The boundary conditions for $r \rightarrow \infty$ are:
\begin{eqnarray}
\bnabla \varPhi &=& -\beta \ui, \\
C &=& 1, \\
\bV &=& -\cU \ui,
\end{eqnarray}
and the total force acting on the particle is zero: $\bF = \bzero$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Asymptotic Analysis} \label{asymptotic}
The nonlinear system can be written as: 
\begin{eqnarray}
\cO(\bx;\beta) &=& \bzero.
\end{eqnarray}

The solution to the non-linear system can be written as a Taylor series in $\beta$.
\begin{eqnarray}
\bx = \bx(\beta) &\approx& \sum_n \bx_n \beta^n.
\end{eqnarray}

All the variables are expanded around $\beta = 0$:
\begin{eqnarray}
\varPhi &=& \beta \varPhi_1 + \beta^2 \varPhi_2 + \beta^3 \varPhi_3 + \ldots, \\
C &=& 1 + \beta C_1 + \beta^2 C_2 + \beta^3 C_3 + \ldots, \\
\Psi &=& \beta \Psi_1 + \beta^2 \Psi_2 + \beta^3 \Psi_3 + \ldots ,
\end{eqnarray}
where the velocity $\bV$ and pressure $P$ are computed from the streamfunction $\Psi$:
\begin{eqnarray}
\bV &=& \bnabla \times \pars{\frac{\Psi}{r \sin\theta} \bphihat} 
= \frac{1}{r^2 \sin\theta} \deriv{\Psi}{\theta} \brhat 
- \frac{1}{r \sin\theta} \deriv{\Psi}{r} \bthetahat.
\end{eqnarray}

The nonlinear equations and the boundary conditions are expanded around $\beta = 0$
and arranged by powers of $\beta$:
\begin{eqnarray}
\bzero = \cO(\bx) = \sum_k \cO_k(\bx_0, \ldots \bx_k) \beta^k
\end{eqnarray}
This can be rewritten as a system of equations:
\begin{eqnarray}
\brc{rcl}{
\bzero &=& \cO_0(\bx_0), \\
\bzero &=& \cO_1(\bx_0, \bx_1), \\
\bzero &=& \cO_2(\bx_0, \bx_1, \bx_2), \\
& \vdots & \\
\bzero &=& \cO_k(\bx_0, \bx_1, \bx_2, \ldots \bx_k).
}
\end{eqnarray}
Thus, the $O(\beta^k)$ term $\bx_k$ can be found recursively by solving 
$\cO_k(\bx_0, \bx_1, \ldots, \bx_k) = \bzero$,
given the previous solutions for $\bx_0, \ldots, \bx_{k-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Linear Solution $(k=1)$}

For $\beta = 0$ (no electric field is applied), the steady-state solution is $\cU = 0$:
\begin{eqnarray}
\varPhi_0(r,\theta) &=& 0, \\
C_0(r,\theta) &=& 1, \\
\bV_0(r,\theta) &=& \bzero, \\
P_0(r,\theta) &=& 0.
\end{eqnarray}

Assuming $\beta \ll 1$, the linearized equations and boundary conditions are 
(as developed in \cite{yariv2010migration}):
\begin{eqnarray}
\Laplacian \varPhi_1 &=& 0, \\
\Laplacian C_1 &=& 0, \\
\bLaplacian \bV_1 - \bnabla P_1 &=& \bzero.
\end{eqnarray}

The first order terms are:
\begin{eqnarray}
\varPhi &=& \beta \pars{\frac{1}{4r^2} - r}\cos\theta, \\
C &=& 1 + \beta \frac{3}{4r^2} \cos\theta, \\
\Psi &=& \beta \cU_1 \pars{\frac{1}{r} - r^2} \frac{\sin^2\theta}{2}, \\
\bV &=& -\beta \cU_1 \brcs{\pars{1 - \frac{1}{r^3}}\cos\theta \cdot \brhat - 
                               \pars{1 + \frac{1}{2r^3}} \sin\theta \cdot \bthetahat}, \\
P &=& 0, \\
\bF &=& \bzero.
\end{eqnarray}

The slip velocity satisfies:
\begin{eqnarray}
\beta \cU_1 \frac{3}{2} \sin\theta =
V_\theta &=& 4 \log \pars{\frac{1 + e ^ \frac{\zeta}{2}}{2}} \deriv{\varPhi}{\theta} 
=
 3 \beta \log \pars{\frac{1 + e ^ \frac{\zeta}{2}}{2}} \sin\theta, \\
\cU_1 &=& 2 \log \pars{\frac{1 + e ^ \frac{\zeta}{2}}{2}} 
       =  2 \log \pars{\frac{1 + \gamma ^ {-\frac{1}{2}}}{2}}.
\end{eqnarray}
because $\varPhi_1 = -C_1 = -\frac{3}{4} \cos\theta$ on $r=1$
and $\zeta_0 = -\log\gamma$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonlinear Solution ($k > 1$)}
Note that, due to symmetry considerations, $\cU(\beta)$ is an anti-symmetric function:
\begin{eqnarray}
\cU(-\beta) &=& -\cU(\beta).
\end{eqnarray}
Therefore, $\cU_2 = 0$ and the next velocity term is the cubic one:
\begin{eqnarray}
\cU(\beta) &\approx& \beta \cU_1 + \beta^3 \cU_3 + O(\beta^5)
\end{eqnarray}

In order to find the cubic terms, the quadratic term should be found first.

\subsubsection{Quadratic Solution ($k = 2$)}
The quadratic terms $\bx_2$ are computed by solving a linear PDE system 
with a right-hand side determined by the linear terms $\bx_1$:
\begin{eqnarray}
\varPhi_2 &=& \frac{\left(\frac{\cU_1\, \alpha}{32} - \frac{1}{16}\right)\, \left(3\, {\cos^2\theta} - 1\right)}{r^3} - \frac{3}{32\, r^4} - \frac{3\, {\sin^2\theta}\, \left(4\, r^3 - 1\right)}{32\, r^4} - \frac{3\, \cU_1\, \alpha - 6}{32\, r},
\\
C_2 &=& \frac{\left(\frac{5\, \cU_1\, \alpha}{32} + \frac{1}{16}\right)\, \left(3\, {\cos^2\theta} - 1\right)}{r^3} - \frac{3\, \cU_1\, \alpha - 6}{32\, r} + \frac{3\, \cU_1\, \alpha\, \left(2\, r^3\, {\sin^2\theta} + {\sin^2\theta} - 1\right)}{16\, r^4},
\\
\Psi_2 &=& \pars{\frac{9}{16(\sqrt{\gamma}+1)} - \frac{3}{16} \cU_1 (\cU_1 \alpha + 1)}
 \left(\frac{1}{r^2} - 1\right) \sin^2\theta \cos\theta,  \\
\bF_2 &=& \bzero, \\ \cU_2 &=& 0.
\end{eqnarray}
Note that the quadratic term does not contribute to the total force and staedy-state 
velocity terms, but it does change the fluid flow, 
the electric potential and the ionic concentration.

\subsubsection{Cubic Solution ($k = 3$)}

\begin{eqnarray*}
\varPhi_3 &=& \cos\theta \left(\frac{15 \cU_1\alpha}{64} + \frac{3}{32}\right) - \frac{3 \cU_1\alpha {\cos}^3\theta}{32} - \frac{\cos\theta \left( - 183 {\cU_1}^2\alpha^2 + 839 \cU_1\alpha + 470\right)}{2560 r^2} \\ 
&&+ \frac{3 \cos\theta \left(5 \cU_1\alpha - 6 {\cos^2\theta} - 4 \cU_1\alpha {\cos^2\theta} + 4\right)}{64 r^3} + \frac{\left(2 \cU_1\alpha - 1\right) \left(\cos\theta - 3 {\cos}^3\theta\right)}{64 r^5} 
\\ 
&&- \frac{\left(3 \cos\theta - 5 {\cos}^3\theta\right) \left( - \frac{19 {\cU_1}^2\alpha^2}{5120} + \frac{97 \cU_1\alpha}{5120} + \frac{3 \cU_2\alpha}{320} + \frac{21}{1280}\right)}{r^4} + \frac{{\cos}^3\theta \left(\frac{15 \cU_1\alpha}{64} + \frac{3}{32}\right)}{r^2} \\ 
&&+ \frac{{\cos}^3\theta \left(\frac{\cU_1\alpha}{64} + \frac{3}{64}\right)}{r^6},
\\
C_3 &=&
\frac{3 {{\cU_1}}^2\alpha^2 {\cos^3\theta}}{32} + \frac{\cos\theta \left(797 {{\cU_1}}^2\alpha^2 + 419 {\cU_1}\alpha - 512 {\cU_2}\alpha + 174\right)}{2560 r^2} \\
&&- \frac{\left(3 \cos\theta - 5 {\cos^3\theta}\right) \left(\frac{59 {{\cU_1}}^2\alpha^2}{5120} + \frac{103 {\cU_1}\alpha}{5120} + \frac{69 {\cU_2}\alpha}{320} + \frac{3}{1280}\right)}{r^4} \\
&&- \frac{3\alpha \cos\theta \left( - 8\alpha {{\cU_1}}^2 {\cos^2\theta} + 7\alpha {{\cU_1}}^2 + 2 {\cU_1} + 32 {\cU_2} {\cos^2\theta} - 32 {\cU_2}\right)}{128 r^3} - \frac{3 {\cU_1}\alpha \cos\theta \left(5 {\cU_1}\alpha + 2\right)}{64} \\
&&+ \frac{\alpha \left(\cos\theta - 3 {\cos^3\theta}\right) \left(5\alpha {{\cU_1}}^2 + 2 {\cU_1} + 16 {\cU_2}\right)}{128 r^5} + \frac{{{\cU_1}}^2\alpha^2 {\cos^3\theta}}{32 r^6} - \frac{3 {\cU_1}\alpha {\cos^3\theta} \left(5 {\cU_1}\alpha + 2\right)}{64 r^2},
\\
\Psi_3 &=&
r^2 {\sin^2\theta} \left(\frac{{\cW_2}}{15} - \frac{{\cW_1}}{3} + \frac{209}{3360}\right) 
 - \frac{{\sin^2\theta} \left(\frac{5 {\cW_2}}{3} - \frac{{\cW_1}}{3} + \frac{35}{264}\right)}{r}
\\ && + \frac{{\sin^4\theta} \left(2 {\cW_2} + \frac{761}{5632}\right)}{r} - \frac{{\sin^2\theta} \left(1848 r^6 {\sin^2\theta} - 2079 r^3 {\sin^2\theta} + 924 r^3 - 7 {\sin^2\theta} + 10\right)}{19712 r^5} 
\\ && + \frac{\left(\frac{2 {\cW_2}}{5} + \frac{829}{28160}\right) \left(4 {\sin^2\theta} - 5 {\sin\theta}^4\right)}{r^3} + \frac{{\sin^2\theta} \left(\frac{{\cU_1}\alpha}{16} - \frac{1}{8}\right) \left(2 r^3 - 3 r^2 + 1\right)}{4 r},
\\
\cW_1 &=& \frac{69}{512 \left(\sqrt{\gamma} + 1\right)} - \frac{1407\alpha^2 {\log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}^3}{1280} - \frac{123 \log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}{1280} - \frac{27}{512 {\left(\sqrt{\gamma} + 1\right)}^2} \\ && -
\frac{1899\alpha {\log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}^2}{2560} - \frac{\alpha \log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right) \left(\frac{81}{16 \left(\sqrt{\gamma} + 1\right)} - \log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right) \left(\frac{27\alpha \log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}{4} + \frac{27}{8}\right)\right)}{320} \\ && + 
\frac{21\alpha \log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}{128 \left(\sqrt{\gamma} + 1\right)},
\\
\cW_2 &=& \frac{9 \log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}{256} + \frac{21\alpha^2 {\log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}^3}{64} - \frac{27}{512 \left(\sqrt{\gamma} + 1\right)} - \frac{27}{512 {\left(\sqrt{\gamma} + 1\right)}^2}
 + \frac{15\alpha {\log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}^2}{64} 
\\ && 
- \frac{297\alpha \log\left(\frac{\sqrt{\gamma} + 1}{2 \sqrt{\gamma}}\right)}{1024 \left(\sqrt{\gamma} + 1\right)}.
\end{eqnarray*}

The cubic solution satisfies the $\lim_{r\rightarrow\infty}C_3 = 0$ boundary condition, 
when $\alpha\cU_1 = 0$.

The steady-state velocity can be expanded using odd powers of $\beta$:
\begin{eqnarray}
\cU(\beta, \gamma) &=& \cU_1(\gamma) \beta + \cU_3(\gamma) \beta^3
\end{eqnarray}

The velocity terms (for $\alpha = 0$) are:

The linear term is:
\begin{eqnarray}
\cU_1(\gamma) &=& 2 \log \pars{\frac{1 + \gamma^{-\frac{1}{2}}}{2}}. \\
\end{eqnarray}
The cubic term is:
\begin{eqnarray}
\cU_3(\gamma) &=& \frac{31}{320(\sqrt\gamma + 1)} - \frac{9}{320(\sqrt\gamma + 1)^2} + \frac{1}{1680} - \frac{11}{160} \log \pars{\frac{1 + \gamma^{-\frac{1}{2}}}{2}}.
\end{eqnarray}

Note that, for $\gamma = 1$, the linear term vanishes and the velocity leading term is $O(\beta^3)$:
\begin{eqnarray}
\cU(\beta) = \frac{1129}{26880}\beta^3.
\end{eqnarray}

Note that a $\cU = 0$ solution may exist for $\beta_c \ne 0$ and $\alpha = 0$, 
if it satisfies:
\begin{eqnarray}
\beta_c^2 = -\frac{\cU_1(\gamma)}{\cU_3(\gamma)}.
\end{eqnarray}
For $\gamma = 1 + \eps$, where $0 < \eps \ll 1$:
\begin{eqnarray}
\beta_c \approx \sqrt{-\frac{\cU_1'(1) \eps}{\cU_3(1)}} = 
 \sqrt{\frac{13440 \eps}{1129}} \approx 3.45 \sqrt{\eps}.
\end{eqnarray}
However, for $\gamma \gg 1$, we have:
\begin{eqnarray}
\beta_c =  \pars{\frac{11}{320} + \frac{1}{1680 \log 4}}^{-\frac{1}{2}} \approx 5.36.
\end{eqnarray}

The cubic term becomes dominant for small $\beta$ when $\gamma \approx 1$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Tests} \label{results}
The results for steady-state velocity $\cU$ as a function of the electric field magnitude $\beta$
are shown at Figures \ref{fig:NumResA} and \ref{fig:NumResB}. 
Since log-log scale is used, blue and red colors are used 
for positive and negative velocities, respectively.

For small $\beta$, the linear regime dominates, $\cU \propto \beta$. 
However, from a specific $\beta_c$, the nonlinear regime dominates, such that
$\cU \propto \beta^3$, until $\beta \sim O(1)$.

The streamlines for a specific $\gamma$ are shown at Figure \ref{fig:NumResC} for
various electric field magnitudes $\beta$.

The theoretical results for the nonlinear regime are computed by asymptotic expansion of the
nonlinear system, as described in the previous section.
\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figs/A1.eps}
        \includegraphics[width=0.45\textwidth]{figs/A2.eps}
        \includegraphics[width=0.45\textwidth]{figs/A3.eps}
        \includegraphics[width=0.45\textwidth]{figs/A4.eps}
        \caption{Steady-state velocity $\cU$ as a function of the electric field magnitude $\beta$, 
        for $\gamma > 1$.}
        \label{fig:NumResA}
    \end{center}
\end{framed}
\end{figure}
\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figs/B1.eps}
        \includegraphics[width=0.45\textwidth]{figs/B2.eps}
        \includegraphics[width=0.45\textwidth]{figs/B3.eps}
        \includegraphics[width=0.45\textwidth]{figs/B4.eps}
        \caption{Steady-state velocity $\cU$ as a function of the electric field magnitude $\beta$,
        for $\gamma < 1$.}
        \label{fig:NumResB}
    \end{center}
\end{framed}
\end{figure}
\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.3\textwidth]{figs/D1n.eps}
        \includegraphics[width=0.3\textwidth]{figs/D1a.eps}

        \includegraphics[width=0.3\textwidth]{figs/D2n.eps}
        \includegraphics[width=0.3\textwidth]{figs/D2a.eps}

        \includegraphics[width=0.3\textwidth]{figs/D3n.eps}
        \includegraphics[width=0.3\textwidth]{figs/D3a.eps}

        \includegraphics[width=0.3\textwidth]{figs/D4n.eps}
        \includegraphics[width=0.3\textwidth]{figs/D4a.eps}
        \caption{Streamlines of the flow for various electric field values: numerical results (left)
        and analytical approximation (right).}
        \label{fig:NumResC}
    \end{center}
\end{framed}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work} \label{conclusions}

\bibliographystyle{elsarticle-num}
\bibliography{main}

\end{document}